---
title: "Gross Square Ft. vs. Sale Price"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Live Session 6 Assignment
```{r}
###plyr library loaded
library(plyr)
### manhattan data read
manhattan <- read.csv("../data/rollingsales_manhattan.csv",skip=4,header=TRUE)
## Check the data
head(manhattan)
summary(manhattan)
str(manhattan)

## clean/format the data with regular expressions
manhattan$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", manhattan$SALE.PRICE))
##count NA values for sale prices in manhattan
count(is.na(manhattan$SALE.PRICE.N))
##lowercase names in manhattan
names(manhattan) <- tolower(names(manhattan)) 
## Get rid of leading digits
manhattan$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", manhattan$gross.square.feet))
manhattan$land.sqft <- as.numeric(gsub("[^[:digit:]]","", manhattan$land.square.feet))
manhattan$year.built <- as.numeric(as.character(manhattan$year.built))

## do a bit of exploration to make sure there's not anything
## weird going on with sale prices
attach(manhattan)
hist(sale.price.n) 
detach(manhattan)
```
###As you can see above, the histogram covers the area far left of $100 million sale price marker. To the right of the first histogram bar, there is a tiny bar which covers the $100-200 millon sale price area.
```{r}

## keep only the actual sales where price is not equal to 0
manhattan.sale <- manhattan[manhattan$sale.price.n!=0,]

## Plot gross square ft vs. sale price for manhattan
plot(manhattan.sale$gross.sqft,manhattan.sale$sale.price.n)

```
##Gross Sq. ft. vs. sale price (log base 10 transformed)
```{r}
## Plot gross square ft vs. sale price for manhattan with log 10 transform
plot(log10(manhattan.sale$gross.sqft),log10(manhattan.sale$sale.price.n))
```
##Gross Sq. ft. vs. sale price (log base 10 transformed) for only 1,2,3 family homes
```{r}

##Filter out family homes from rest of manhattan data
manhattan.homes <- manhattan.sale[which(grepl("FAMILY",manhattan.sale$building.class.category)),]
## take dimension of this new data frame
dim(manhattan.homes)

## plot log 10 transform of family homes gross square ft vs. sale price
plot(log10(manhattan.homes$gross.sqft),log10(manhattan.homes$sale.price.n))
summary(manhattan.homes[which(manhattan.homes$sale.price.n<100000),])
```
##Gross Sq. ft. vs. sale price(with removed outliers)
```{r}
#outliers when log 10 transform is less than or equal to 5
manhattan.homes$outliers <- (log10(manhattan.homes$sale.price.n) <=5) + 0
#filter out outliers
manhattan.homes <- manhattan.homes[which(manhattan.homes$outliers==0),]

#plot data without outliers
plot(log10(manhattan.homes$gross.sqft),log10(manhattan.homes$sale.price.n))
```
